---
title: "Data Analysis Project"
author: "Kassidy, Maddie, Cassidie"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
options(scipen=999)
library(tidyverse)
library(lubridate)
library(refinr)
library(tidycensus)
install.packages("censusxy")
library(censusxy)

library(janitor)
library(sf)
library(tigris)
install.packages("usmap")
library(usmap)

census_api_key("0105753aba54d204fb6c33c44044ae59449b3773", overwrite =TRUE)


```

## Load and Cleaning Data

```{r}
# Load geocoded shootings data

fatal_shootings <- read_rds("data/all_geocoded_shootings.rds")

shootings_geoid_grouped <- fatal_shootings %>%
  group_by(census_tracts_geoid)%>%
  summarise(count=n())%>%
  rename(geoid=census_tracts_geoid)

# # Load census data
# states_dataframe <- fips_codes %>%
#   distinct(state) %>%
#   head(51) %>%
#   as_vector()

# census <- get_acs(geography = "tract", state=states_dataframe, 
#                        variables = c(popululation = "B02001_001",
#                                 median.gross.rent = "B25064_001",
#                                 median.household.income = "B19013_001",
#                                 rent.burden = "B25071_001",
#                                 white = "B03002_003", 
#                                 af.am = "B03002_004",
#                                 hispanic = "B03002_012",
#                                 am.ind = "B03002_005",
#                                 asian = "B03002_006",
#                                 nh.pi = "B03002_007",
#                                 multiple = "B03002_009",
#                                 other = "B03002_008"),
#                           year= 2019)
#write_rds(census,"census.rds")
census <- read_rds("data/census.rds")%>%
  clean_names()%>%
  select(-moe) %>%
  pivot_wider(names_from=variable, values_from=estimate) %>%
  rename(population = popululation) %>%
  mutate(percentage_black = (`af.am`/population))

glimpse(census)

shootings_census <- census %>%
  left_join(shootings_geoid_grouped) %>%
  mutate(shooting=case_when(
    is.na(count)~"No",
    TRUE~"Yes"
  ))%>%
  select(-count)



```

###Question 1:
Are Police shootings primarily only happening in majority Black neighborhoods even when the victim is not black?
###Answer 1:
Yes, it is happening more predominately in Black neighborhoods, the percentage overall in the US for the black population is 12%, and the shooting data showed it as 14% in Black areas.

```{r}

shootings_census %>%
  filter(shooting=="Yes") %>%
  summarise(total_black = sum(af.am),
           total_pop = sum(population)) %>%
  mutate(percentage_black=(total_black/total_pop))


shootings_census %>%
  summarise(total_black = sum(af.am),
           total_pop = sum(population)) %>%
  mutate(percentage_black=(total_black/total_pop))


#By Census tract

sc <- shootings_census %>%
  filter(shooting=="Yes") %>%
  filter(percentage_black >= .5) %>%
  arrange(desc(percentage_black))
  

```

### Question 2
*Do these shootings mostly occur in low income areas? 
###Answer2:
Yes, shootings are occurring more in low income areas. The mean income in the shooting data is $56,554, where the national mean income is $67,007.

```{r}
#Aggregate
shootings_income<-shootings_census%>%
  filter(!is.na(median.household.income)) %>%
  summarise(mean_income =mean(median.household.income))


census_income<-census%>%
  filter(!is.na(median.household.income)) %>%
  summarise(mean_income =mean(median.household.income))





```

### Question 3
*Are police shootings still happening disproportionately affecting black people regardless of the area (even in white neighborhoods)?
###Answer 3:
Yes, police shootings are disproportionately affecting Black people. They account for 23% of police shootings, even though they account for 12% of the country's entire population. 


```{r}
options(scipen = 999)

race_shootings <- shootings_census %>%
  filter(!is.na(race))

race_counted<-race_shootings%>%
 group_by(race)%>%
  count(race)

race<-race_counted%>%
  add_column(total="6410")

race$total <- as.numeric(race$total)

race_percent<-race%>%
   mutate(percent=(n/total))
```

### Question 4
How do these percentages compare to other states.
### Answer 4
In Illinois, the Black population accounts for about 14% of the total population, yet in 50% percent of its shootings the victim is Black. 

```{r}
#Filtering for only Black shootings
shootings_geoid_black <- shootings_census %>%
  filter(race =="B")

#Grouping Black shootings by state
shootings_geoid_black <- shootings_geoid_black %>%
  group_by(state) %>%
  summarise(count=n()) %>%
  arrange(desc(count))

#Total shootings per state
shootings_states <- shootings_census %>%
  group_by(state) %>%
  summarise(count=n()) %>%
  arrange(desc(count)) %>%
  rename(total_count=count)

#Joining by state
shootings_total <- shootings_states %>%
  left_join(shootings_geoid_black, by=c("state"))

#Find percentage
shootings_total<-shootings_total %>%
  mutate(percentage_black=(count/total_count)) %>%
  arrange(desc(percentage_black))

#Getting Census data for population by state
census_2 <- get_acs(geography = "state", 
                       variables = c(popululation = "B02001_001",
                                median.gross.rent = "B25064_001",
                                median.household.income = "B19013_001",
                                af.am = "B03002_004"),
                          year= 2019)

write_rds(census_2,"census_2.rds")
census_2 <- read_rds("census_2.rds")

census_3 <- census_2 %>%
  select(-moe) %>%
  pivot_wider(names_from=variable, values_from=estimate)

census_3 <- census_3 %>%
  mutate(total_pop_black=(af.am/popululation)) %>%
  arrange(desc(total_pop_black)) %>%
  rename(state_name=NAME)

fips_codes <- fips_codes %>%
  select(state,state_name)%>%
  distinct()

shootings_black_joined <- fips_codes %>%
  left_join(shootings_total, by=c("state"))%>%
  left_join(census_3, by=c("state_name")) %>%
  mutate(difference=(percentage_black-total_pop_black)) %>%
  arrange(desc(difference))


#Illinois

race_shootings %>%
  group_by(city)%>%
  filter(state=="IL")
  

```
### Final Memo
For our final data analysis project, our group chose to find out more about police shootings and whether race, income level, and location play a part in whether these fatal shootings occur or not. We took data from the Washington Post’s police shootings database, which is extensive and well-updated regularly. We found through our analysis that black people are disproportionately affected by fatal police shootings, that fatal police shootings mostly occur in low-income areas with high percentages of residents being black, and that in states like Illinois where the black population only makes up for 14% of the population, 50% of the time there is a fatal shooting the victim is black. 
We first found that police shootings are disproportionately affecting black people, regardless of their location. The black population accounts for 12% of the country’s total population, yet they account for 23% of victims in police shootings. In states like Illinois, the black population accounts for 14% of the total state population. But, the black population also accounts for half of the victims in police shootings in this state. 
Throughout our analysis, we found that police shootings primarily happen in black neighborhoods, regardless of the victim’s race.
We then used the data to find whether or not there was a high number of police shootings where there was a majority black population. To do this we filtered the shootings data to include any instances where more than 50% of the population was black based on the area they lived in, which we got from the census data. We decided that “majority-black” could be any census tract where 50% or more residents were African American. This analysis showed us that 14% of areas where police shootings occur have a population where over half the residents are black. We also found that lower median household income correlated pretty closely with higher percentage black populations. For example, Orleans Parish, Louisiana had the lowest median household income in our data, and 85% of its population was black. We found that in our data frame there were only 31 census tracts that had median income levels that were over the national mean income, out of a total of 626 census tracts that had over 50% of their population is black. 
We then found that police shootings occur more frequently in low-income areas. The mean income within the areas where police shootings have occurred is $56,554. This is below the national mean income of $67,007 across the country.
The police shootings data required little cleaning and consisted of detailed information about the shooting, such as the victim’s personal information and location of the shooting. The data also included information like the date the victim was shot, if they were armed, if there were signs of mental illness in the victim, and information on whether they tried to flee. The difficult aspect of the data was that it did not include the census tracts GEOID. We analyzed this data with API census data in order to gain a better understanding of where these shootings are occurring and the demographics of these areas. It became difficult to organize that data after joining the census tracts GEOID with the original data frame. We also found another difficult aspect of this data was that we needed to make a lot of separate data frames and join different aspects of each one to create the information we needed to answer our questions. 
In order to verify our findings, we found areas in which proved our findings to be true. For example, after determining that police shootings disproportionately affect black people, we then used our original police shootings data frame to see in what states this statement is true. We filtered the data to observe the percentage of black victims that are shot by police per state. Then we used Census API data to determine the percentage in which the black population accounts for out of the entire state population. We found that in Illinois, the black population accounts for 14% of the state’s total population, yet in 50% of police shootings in the state, the victim is black. This verifies that police shootings disproportionately affect black people. 
The steps we would take to finish reporting this story would include focusing on census tracts that have the most disproportionality in terms of people being shot by law enforcement officers in predominantly black neighborhoods. This would include places like Illinois as we noted previously. In Illinois, we could further use the data to report out where these shootings are occurring in the state. This could possibly lead us to understand why black people account for so little of the population yet account for half of police shootings. We could also look at Orleans Parish, Louisiana, which we also previously talked about because of its extremely low median household income and the percentage of its black residents where shootings have occurred. To make the story more real to readers, we would need to do additional reporting. An idea we had would be to interview people who have been affected by these police shootings in predominantly black low-income areas. Highlighting one specific story could add a lot of emotion to a story that has a lot of data and numbers. We believe that in focusing on a place like Orleans Parish, we could possibly interview residents who fit the criteria of our findings, which would be someone who has been affected by police shootings who have lived in this particular census tract. Focusing on one story in particular where someone has lost a loved one due to a fatal police shooting would add something to our story that data could not provide for the reader. In addition to this, interviewing the local police departments in these areas, such as the New Orleans Police Department and the Orleans Parish Sheriff’s Department, could be useful when reporting these findings. It would be interesting to see if government employees, such as police officers, and government officials are aware of this disparity in their communities. 




-30-
