---
title: "Data Analysis Project"
author: "Kassidy, Maddie, Cassidie"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
options(scipen=999)
library(tidyverse)
library(lubridate)
library(refinr)
```

## Load and Cleaning Data

```{r}
# Load required data
fatal_shootings<- read_csv("data/fatal-police-shootings-data.csv") 
# Clean required data and prepare for analysis if needed. 
```



### Question 1
*Are police shootings happening mostly in majority-black communities, even when the victim is not black?

```{r}
```

### Question 2
*Do these shootings mostly occur in low income areas? 

```{r}
```

### Question 3
*Are police shootings still happening disproportionately affecting black people regardless of the area (even in white neighborhoods)?

```{r}
```


-30-
