---
title: "Reverse Engineering Project"
author: "Maddie Osgood, Kassidy McDonald, Cassidie Stevens"
date: "October 6, 2021"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are reverse engineering the story, [As police struggle to solve homicides, Baltimore residents see an 'open season for killing'](https://www.washingtonpost.com/investigations/as-police-struggle-to-solve-homicides-baltimore-residents-see-an-open-season-for-killing/2018/12/26/7ee561e4-fb24-11e8-8c9a-860ce2a8148f_story.html)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
options(scipen=999)
library(tidyverse)
library(lubridate)
library(refinr)
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

```{r}
# Load required data
baltimore_homicide <- read_csv("data/homicide-data.csv")
# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")

# Clean required data and prepare for analysis if needed. 


```

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

* **Sentence text**: City police made an arrest in 41 percent of homicides in 2014; last year (2017), the rate was just 27 percent, a 14 percentage point drop.
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]
After the results of our analysis, the statistic that city police made an arrest in 41 percent of homicides in 2014 is correct. The statistic that police made an arrest in 27 percent of homicides in 2017 is also correct. Lastly, by combining the 2014 homicide percentage and the 2017 percentage data frames and then subtracting the two percentages from each other, we confirmed that it was a 14 percentage point drop between 2014 and 2017.

```{r}
#2014
percent_2014 <- baltimore_homicide %>% 
  filter(city == "Baltimore") %>%
  mutate(reported_date = as.Date(as.character( reported_date), format="%Y%m%d")) %>%
  filter(reported_date >= as.Date("2014-01-01"), reported_date <= as.Date("2014-12-31")) %>%
  group_by(disposition)%>%
   summarise(
    count= n()
  ) %>%
  mutate(total_homicides = sum(count)) %>%
  filter(disposition == "Closed by arrest") %>%
  mutate(percentage2014 = count / total_homicides)

#2017
percent_2017 <- baltimore_homicide %>% 
  filter(city == "Baltimore") %>%
  mutate(reported_date = as.Date(as.character( reported_date), format="%Y%m%d")) %>%
  filter(reported_date >= as.Date("2017-01-01"), reported_date <= as.Date("2017-12-31")) %>%
  group_by(disposition)%>%
   summarise(
    count= n()
  ) %>%
  mutate(total_homicides = sum(count)) %>%
  filter(disposition == "Closed by arrest") %>%
  mutate(percentage2017 = count / total_homicides) 
  
#difference
total_2014_2017 <- bind_cols(percent_2014,percent_2017)%>%
  mutate(percentage2017-percentage2014)

```

### Sentence 2

* **Sentence text**:In Chicago, the homicide arrest rate has dropped 21 percentage points, in Boston it has dropped 12 points and in St. Louis it is down 9. (2014 vs 2017)
* **Analysis summary**: Through filtering the data to only homicides that resulted in an arrest in 2014 and 2015 in Chicago, we subtracted the percent of total homicides from each year to confirm that the homicide arrest rate in Chicago has dropped 21 percentage points. We did this same process for the cities of Boston and St. Louis to confirm that Boston dropped 12 percentage points and St. Louis dropped 9 percentage points.

```{r}
# Put code to reverse engineer sentence here

#Chicago

#Chicago 2014
chicago_percent_2014<- baltimore_homicide %>% 
  filter(city == "Chicago") %>%
  mutate(reported_date = as.Date(as.character( reported_date), format="%Y%m%d")) %>%
  filter(reported_date >= as.Date("2014-01-01"), reported_date <= as.Date("2014-12-31")) %>%
group_by(disposition)%>%
   summarise(
    count= n()
  ) %>%
  mutate(total_homicides = sum(count)) %>%
  filter(disposition == "Closed by arrest") %>%
  mutate(percentagechicago2014 = count / total_homicides) 
  
#Chicago 2017
chicago_percent_2017<- baltimore_homicide %>% 
  filter(city == "Chicago") %>%
  mutate(reported_date = as.Date(as.character( reported_date), format="%Y%m%d")) %>%
  filter(reported_date >= as.Date("2017-01-01"), reported_date <= as.Date("2017-12-31")) %>%
group_by(disposition)%>%
   summarise(
    count= n()
  ) %>%
  mutate(total_homicides = sum(count)) %>%
  filter(disposition == "Closed by arrest") %>%
  mutate(percentagechicago2017 = count / total_homicides) 

#CHICAGO PERCENT CHANGE
chicago_total <- bind_cols(chicago_percent_2014,chicago_percent_2017)%>%
  mutate(change = percentagechicago2017-percentagechicago2014)

#Boston
#Boston 2014
boston_2014 <- baltimore_homicide %>% 
  filter(city == "Boston") %>%
  mutate(reported_date = as.Date(as.character( reported_date), format="%Y%m%d")) %>%
  filter(reported_date >= as.Date("2014-01-01"), reported_date <= as.Date("2014-12-31")) %>%
group_by(disposition)%>%
   summarise(
    count= n()
  ) %>%
  mutate(total_homicides = sum(count)) %>%
  filter(disposition == "Closed by arrest") %>%
  mutate(percentageboston2014 = count / total_homicides)


#Boston 2017
boston_2017<- baltimore_homicide %>% 
  filter(city == "Boston") %>%
  mutate(reported_date = as.Date(as.character( reported_date), format="%Y%m%d")) %>%
  filter(reported_date >= as.Date("2017-01-01"), reported_date <= as.Date("2017-12-31")) %>%
group_by(disposition)%>%
   summarise(
    count= n()
  ) %>%
  mutate(total_homicides = sum(count)) %>%
  filter(disposition == "Closed by arrest") %>%
  mutate(percentageboston2017 = count / total_homicides) 
  
#Boston Percent Change
boston_total <- bind_cols(boston_2014,boston_2017)%>%
  mutate(percentageboston2017-percentageboston2014)
  
#St. Louis

#St. Louis 2014
St_Louis_2014 <- baltimore_homicide %>% 
  filter(city == "St. Louis") %>%
  mutate(reported_date = as.Date(as.character( reported_date), format="%Y%m%d")) %>%
 filter(reported_date >= as.Date("2014-01-01"), reported_date <= as.Date("2014-12-31")) %>%
group_by(disposition)%>%
   summarise(
    count= n()
  ) %>%
  mutate(total_homicides = sum(count)) %>%
  filter(disposition == "Closed by arrest") %>%
  mutate(percentage_stlouis_2014 = count / total_homicides)
 
 
#St. Louis 2017
St_Louis_2017 <- baltimore_homicide %>% 
  filter(city == "St. Louis") %>%
  mutate(reported_date = as.Date(as.character( reported_date), format="%Y%m%d")) %>%
 filter(reported_date >= as.Date("2017-01-01"), reported_date <= as.Date("2017-12-31")) %>%
group_by(disposition)%>%
   summarise(
    count= n()
  ) %>%
  mutate(total_homicides = sum(count)) %>%
  filter(disposition == "Closed by arrest") %>%
  mutate(percentage_stlouis_2017 = count / total_homicides)

#St. Louis Percent Change
St_Louis_total <- bind_cols(St_Louis_2014,St_Louis_2017)%>%
  mutate(percentage_stlouis_2017-percentage_stlouis_2014)
  

# Display results of code below this codeblock

```

### Sentence 3

* **Sentence text**:  Of the 1,002 homicides between 2015 and the beginning of this year, just 252 — one out of every four — resulted in an arrest. 
* **Analysis summary**: Through filtering the data by homicides between the beginning of 2015 to the end of 2017, we can confirm that there were 1,002 homicides in this time period. By grouping the data by dispositions listed as "closed by arrest", we can confirm that 252 cases resulted in arrest. By finding the percent of cases resulting in arrest by the total homicide cases, we can confirm that one out of ever four case resulted in an arrest.

```{r}
# Put code to reverse engineer sentence here
baltimore_homicide %>% 
  filter(city == "Baltimore") %>%
  mutate(reported_date = as.Date(as.character( reported_date), format="%Y%m%d")) %>%
  filter(reported_date >= as.Date("2015-01-01"), reported_date <= as.Date("2017-12-31")) %>%
group_by(disposition)%>%
   summarise(
    count= n()
  ) %>%
  mutate(total_homicides = sum(count)) %>%
  filter(disposition == "Closed by arrest") %>%
  mutate(percentage = count / total_homicides) 

# Display results of code below this codeblock

```

### Sentence 4

* **Sentence text**: Baltimore is also one of 30 cities that have seen an increase in homicides in recent years, with the greatest raw number increase in killings of any city other than Chicago, which has four times the population.
* **Analysis summary**: After creating a data frame that compares the difference between homicides in 2014 and 2017 per city, the statement made by the Washington Post is incorrect. Baltimore is one of 25 cities that have seen an increase in homicides in recent years. The second statement is correct. Baltimore has the greatest raw number increase in killings, 129, of any city other than Chicago, which has an increase of 236 homicides. The data does not have information concerning the population of each city, but by Google search, Baltimore has a population of approximately 609,032 and Chicago has a population of approximately 2.71 million. Chicago has 4.45 times the population as Baltimore, which can be rounded down to four times the population. Therefore, the last statement is correct as well. 

```{r}
#Code that organizes top 30 cities (2014 v 2017)
homicide_2014 <- baltimore_homicide %>%
  mutate(reported_date = as.Date(as.character( reported_date), 
  format="%Y%m%d")) %>%
  filter(reported_date >= as.Date("2014-01-01"), reported_date <= as.Date("2014-12-31")) %>%
  group_by(city) %>%
  summarise(
    count= n()) %>%
   arrange(desc(count))


homicide_2017 <- baltimore_homicide %>%
  mutate(reported_date = as.Date(as.character( reported_date), 
  format="%Y%m%d")) %>%
  filter(reported_date >= as.Date("2017-01-01"), reported_date <= as.Date("2017-12-31")) %>%
  group_by(city) %>%
  summarise(
    count= n()) %>%
   arrange(desc(count)) %>%
  rename(count2 = count)

alldata <- homicide_2014 %>%
  inner_join(homicide_2017)

alldata <- alldata %>%
  mutate(difference=count2 - count) %>%
  arrange(desc(difference))


  

```

Final Memo:
  The main findings of this story by the Washington Post are that data shows that homicides in Baltimore are increasing, but arrests made by Baltimore police for these homicides are decreasing. The Post notes that this increase in violence began not long after the tragic death of Freddie Gray in 2015, a young man who died while in police custody, sparking an outcry across the country. Data analysis showed that before 2015, Baltimore’s annual homicide arrest rate stayed around 40 percent. The drop of 15 percentage points after Gray's death was sudden, and the Post said it was “the largest single-year drop for any city already solving less than half its homicides.” 
	Another main finding of the story concluded that on average, each Baltimore detective is currently responsible for about 9 homicide cases, and when adding other cases, the total number of cases reach 31. When the Post analyzed police departments that have successful rates of arrests, they found that they usually assign their detectives fewer than 5 cases per year. 
	One last main finding had to do with the city’s black “butterfly,” which is a set of neighborhoods in Baltimore where homicide rates have soared. In each of these “butterfly” neighborhoods,  police make an arrest for only 25 percent of the homicide cases. The Post noted that these are some of the most economically depressed areas, populated almost exclusively by low-income black residents. They also noted how this problem directly relates to the distrust between communities and police officers. The relationship has never been “good,” making it hard for police to find witnesses to help them make arrests, but it has become more distrusting since Gray’s death in 2015.
	Besides using data to show the effect of homicides increasing and arrests decreasing, reporters made this piece come to life by using many victims/their families' stories. Daphne Alston was one example of this. Her son, Tariq Alston, was murdered in 2008. Daphne helped found the Mothers of Murdered Sons and Daughters United, and tells her story to the Post of how she used to attend every funeral, but with the recent spike there’s now too many for her to attend. Her story, along with others included like Cynthia Bruce, who’s son Marcus Tafari Samuel Downer was shot and killed in Baltimore in 2015, add an element of emotion to this piece. When presented with the data along with the personal stories, the piece is effective in showing the reader how devastating this problem is for victims, families, and communities. 
	The reporters communicated their specific data findings in the piece by incorporating facts into the story. We thought some of the most important included: “City police made an arrest in 41 percent of homicides in 2014; last year (2017), the rate was just 27 percent, a 14 percentage point drop,” “In Chicago, the homicide arrest rate has dropped 21 percentage points, in Boston it has dropped 12 points and in St. Louis it is down 9. (2014 vs 2017),” “Of the 1,002 homicides between 2015 and the beginning of this year, just 252 — one out of every four — resulted in an arrest,” and “Baltimore is also one of 30 cities that have seen an increase in homicides in recent years, with the greatest raw number increase in killings of any city other than Chicago, which has four times the population.” We thought that the reporters did an effective job of including their findings into the story in these sentences because they made it simple enough for the average reader to grasp. The database of homicides and amount of data that the reporters had to collect was a lot of information, and by using the strategy of storytelling while also using data makes it easier and more interesting for the reader.
  The Post has an interactive database of homicides where you can see specific statistics in 50 cities in the U.S. The Post was able to do this by collecting data on more than 52,000 homicides in 50 of the largest U.S. cities from the past decade. The data collected included location, arrest status, and information about the victim. Reporters did have to clean and standardize the data they receive from police departments all over the country. They also had to fill in the gaps with some data because police departments only sent them partial information about the homicides. They often had to look up death certificates, court records, and medical examiner reports to find some of the information that they needed.  

-30-