---
title: "Reverse Engineering Project"
author: "Maddie Osgood, Kassidy McDonald, Cassidie Stevens"
date: "October 6, 2021"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are reverse engineering the story, [As police struggle to solve homicides, Baltimore residents see an 'open season for killing'](https://www.washingtonpost.com/investigations/as-police-struggle-to-solve-homicides-baltimore-residents-see-an-open-season-for-killing/2018/12/26/7ee561e4-fb24-11e8-8c9a-860ce2a8148f_story.html)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
options(scipen=999)
library(tidyverse)
library(lubridate)
library(refinr)
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

```{r}
# Load required data
baltimore_homicide <- read_csv("data/homicide-data.csv")
# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")

# Clean required data and prepare for analysis if needed. 


```

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

* **Sentence text**: City police made an arrest in 41 percent of homicides in 2014; last year (2018), the rate was just 27 percent, a 14 percentage point drop.
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]
After the results of our analysis, the statistic that city police made an arrest in 41 percent of homicides in 2014 is correct. We attempted to analyze the 2018 statistic, only to find that the data frame doesn't include any data from 2018. 

```{r}
#2014
baltimore_homicide %>% 
  filter(city == "Baltimore") %>%
  mutate(reported_date = as.Date(as.character( reported_date), format="%Y%m%d")) %>%
  filter(reported_date >= as.Date("2014-01-01"), reported_date <= as.Date("2014-12-31")) %>%
  group_by(disposition)%>%
   summarise(
    count= n()
  ) %>%
  mutate(s = sum(count)) %>%
  filter(disposition == "Closed by arrest") %>%
  mutate(c = sum(count)) %>%
  mutate(percentage = c / s) %>%
  select(percentage) %>%
  group_by(percentage) %>%
  slice(1:1)

#2018
baltimore_homicide %>% 
  filter(city == "Baltimore") %>%
  mutate(reported_date = as.Date(as.character( reported_date), format="%Y%m%d")) %>%
  filter(reported_date >= as.Date("2018-01-01"), reported_date <= as.Date("2018-12-31")) %>%
  group_by(disposition)%>%
   summarise(
    count= n()
  ) %>%
  mutate(s = sum(count)) %>%
  filter(disposition == "Closed by arrest") %>%
  mutate(c = sum(count)) %>%
  mutate(percentage = c / s) %>%
  select(percentage) %>%
  group_by(percentage) %>%
  slice(1:1)

```

### Sentence 2

* **Sentence text**: Of the 1,002 homicides between 2015 and the beginning of this year, just 252 — one out of every four — resulted in an arrest. In Chicago, the homicide arrest rate has dropped 21 percentage points, in Boston it has dropped 12 points and in St. Louis it is down 9.
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 3

* **Sentence text**:  Of the 1,002 homicides between 2015 and the beginning of this year, just 252 — one out of every four — resulted in an arrest. 
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 4

* **Sentence text**: Baltimore is also one of 30 cities that have seen an increase in homicides in recent years, with the greatest raw number increase in killings of any city other than Chicago, which has four times the population.
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 5

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

-30-